### 2020-01-04 figures/finite-classification.tex
lots

\begin{figure}[h]
  \centering
  \begin{tikzcd}[cramped, row sep=small]
    &
    \ar[ddddd, dash, start anchor={[yshift=4ex]}, end anchor={[yshift=-4ex]}] &
    \text{Ordered} &
    \ar[ddddd, dash, dashed, start anchor={[yshift=4ex]}, end anchor={[yshift=-4ex]}] &
    \text{Unordered}
    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \ar[rrrr, dash, start anchor={[xshift=-6ex]}, end anchor={[xshift=10ex]}] &
    &
    &
    &
    {}
    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    &
    &
    \text{Split Enumerable}
      \; \ref{split-enumerable}
      \ar[d, xshift=-1ex, bend right=30]
      &
    &
    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \text{Discrete} &
    &
    \text{Manifest Bishop}
      \ar[u, xshift=1ex, bend right=30]
      \ar[rr, end anchor=south west, start anchor=south east, bend right=10, crossing over]
      \; \ref{manifest-bishop}
      &
    &
    \text{Cardinal}
      \; \ref{cardinal}
      \ar[ll, start anchor=north west, end anchor=north east, bend right=10, "\text{linOrd}" description]
    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \ar[rrrr, dash, dashed, start anchor={[xshift=-6ex]}, end anchor={[xshift=10ex]}] &
    &
    &
    &
    {}
    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \text{Non Discrete} &
    {} &
    \text{Manifest Enumerable}
      \; \ref{manifest-enumerability}
      \ar[from=uu, xshift=-1ex, bend right=30, crossing over]
      \ar[uu, bend right=30,  "\text{Discrete}" description]
      \ar[rr, start anchor=south east, end anchor=south west, bend right=10, crossing over]
      &
    {}
    &
    \text{Kuratowski}
      \; \ref{kuratowski}
      \ar[ll, start anchor=north west, end anchor=north east, bend right=10,  "\text{linOrd}" description]
      \ar[uu, xshift=1ex, bend right=30,  "\text{Discrete}" description]
      \ar[from=uu, xshift=-1ex, bend right=30, crossing over]
  \end{tikzcd}
  \caption{Classification of the strong definitions of finiteness, according to
    whether they are discrete (imply decidable equality) and whether they induce
    a linear order.}
  \label{finite-classification}
\end{figure}