### 2020-05-29 paper.tex
split into sections

\input{preamble}

\author{
  Donnacha Oisín Kidney \and
  Gregory Provan \and
  Nicolas Wu
}
% \affiliation{University College Cork}
% \email{o.kidney@cs.ucc.ie}
% \author{Gregory Provan}
% \affiliation{University College Cork}
% \email{g.provan@cs.ucc.ie}
% \author{Nicolas Wu}
% \affiliation{Imperial College London}
% \email{n.wu@imperial.ac.uk}

\title{Finiteness in Cubical Type Theory}

% \keywords{Agda, Homotopy Type Theory, Cubical Type Theory,
%   Dependent Types, Finiteness, Topos, Kuratowski finite}

\begin{document}

\maketitle
% \begin{abstract}
%   We study five different notions of finiteness in Cubical Type Theory and prove
%   the relationship between them.
%   In particular we show that any totally ordered Kuratowski finite type is
%   manifestly Bishop finite.

%   We also prove closure properties for each finite type, and classify them
%   topos-theoretically.
%   This includes a proof that the category of decidable Kuratowski finite sets
%   (also called the category of cardinal finite sets) form a \(\Pi\)-pretopos.

%   We then develop a parallel classification for the countably infinite types, as
%   well as a proof of the countability of \(A^\star\) for a countable type \(A\).

%   We formalise our work in Cubical Agda, where we implement a library for proof
%   search (including combinators for level-polymorphic fully generic currying).
%   Through this library we demonstrate a number of uses for the computational
%   content of the univalence axiom, including searching for and synthesising
%   functions.
% \end{abstract}
\include{sections/introduction}
\include{sections/finiteness-predicates}
\include{sections/topos}
\include{sections/countable-predicates}
\include{sections/search}
\include{sections/related-work}

% \begin{acks}
%   This work has been supported by the Science Foundation Ireland under the
%   following grant: 13/RC/2D94 to Irish Software Research Centre.
% \end{acks}

\bibliographystyle{plain}
\bibliography{bibliography}
\end{document}