### 2020-06-07 sections/countdown.tex
Update countdown.tex

\chapter{Countdown}
Countdown is a well-known functional programming puzzle (with a spin-off TV show
in the UK), first popularised as a puzzle in which to demonstrate functional
algorithms in \cite{huttonCountdownProblem2002}.
The idea is simple: given a list of numbers, contestants must construct an
arithmetic expression (using a small set of functions) using some or all of the
numbers, to reach some target.

Take the following problem as an example:
\begin{gather*}
  \boxed{1} \boxed{3} \boxed{7} \boxed{10} \boxed{25} \boxed{50} \\
  \boxed{765} \tag{Target}
\end{gather*}
It has the following answer:
\begin{equation}
  3 \times ((7 \times (50 - 10)) - 25)
\end{equation}
Importantly, we do not have to use every number given to get to the target.
For our problem, we will use the operators \(\times\), \(+\), and \(-\).

In this section we will develop a program/proof which can decide countdown
problems totally.
In other words, given a list of numbers and a target, our program will prove
whether a solution exists, and if so, it will provide just that solution.
\section{Classifying The Problem}
So what is a ``solution'' to the countdown problem?
Put simply, it is a way to:
\begin{enumerate}
  \item Arrange some or all of the given numbers into a valid expression
  \item Such that the expression, when evaluated, is equal to the target.
\end{enumerate}

What is the type of the following transformation?

Firstly, we can notice that not every number was used in our solution.
So the transformation must information about what to keep and what to discard on
an item-by-item basis: a subsequence, in other words.

\begin{figure}
  \tikzset{>=Triangle[open]}
  \begin{tikzpicture}
    \node (o1)  at (-1, 1.5) { 1} ;
    \node (o3)  at ( 0, 1.5) { 3} ;
    \node (o7)  at ( 1, 1.5) { 7} ;
    \node (o10) at ( 2, 1.5) {10} ;
    \node (o25) at ( 3, 1.5) {25} ;
    \node (o50) at ( 4, 1.5) {50} ;

    \node(t1) at (-1,  0) {\xcancel{1}}  ;
    \node(t3)   at (0 ,  0) {3}  ;
    \node(t7)   at (1 ,  0) {7}  ;
    \node(t10)  at (2 ,  0) {10} ;
    \node(t25)  at (3 ,  0) {25} ;
    \node(t50)  at (4 ,  0) {50} ;

    \draw [-{Rays[n=4]}] (o1.south)  -- (t1.north) ;
    \draw [->] (o3.south)  -- (t3.north) ;
    \draw [->] (o7.south)  -- (t7.north) ;
    \draw [->] (o10.south) -- (t10.north) ;
    \draw [->] (o25.south) -- (t25.north) ;
    \draw [->] (o50.south) -- (t50.north) ;

    \node(b3)   at (0 , -1.5) {3}  ;
    \node(b7)   at (1 , -1.5) {7}  ;
    \node(b50)  at (2 , -1.5) {50} ;
    \node(b10)  at (3 , -1.5) {10} ;
    \node(b25)  at (4 , -1.5) {25} ;

    \draw [->, rounded corners] (t3.south) --  (b3.north) ;
    \draw [->, rounded corners] (t7.south) --  (b7.north) ;
    \draw [->, rounded corners] (t10.south) to[out=-90, in=90] (b10.north) ;
    \draw [->, rounded corners] (t25.south) to[out=-90, in=90] (b25.north) ;
    \draw [->, rounded corners] (t50.south) to[out=-90, in=90] (b50.north) ;

    \node[draw] at (0.5 , -2.25) {$\times$} ;
    \node[draw] at (1.5 , -2.25) {$\times$} ;
    \node[draw] at (2.5 , -2.25) {$-$} ;
    \node[draw] at (3.5 , -2.25) {$-$} ;

    \node at (0.5 , -3) {$\times$} ;
    \node at (1.5 , -3) {$\times$} ;
    \node at (2.5 , -3) {$-$} ;
    \node at (3.5 , -3) {$-$} ;

    \node(e3) at (0   , -3) {3}  ;
    \node(e7) at (1   , -3) {7}  ;
    \node(e50) at (2   , -3) {50} ;
    \node(e10) at (3   , -3) {10} ;
    \node(e25) at (4   , -3) {25} ;

    \draw[-] (e50.south) -- ++(0.5,-0.75) ;
    \draw[-] (e10.south) -- ++(-1 ,-1.5  ) ;
    \draw[-] (e7.south)  -- ++(1.5,-2.25) ;
    \draw[-] (e3.south)  -- ++(2  , -3 ) ;
    \draw[-] (e25.south) -- ++(-2 , -3 ) ;

    % \node(r40)  [circle, inner sep=0pt, fill=white, anchor=center] at (2.5 , -3.75) {40} ;
    % \node(r280) [circle, inner sep=0pt, fill=white, anchor=center] at (2   , -4.5) {280} ;
    % \node(r255) [circle, inner sep=0pt, fill=white, anchor=center] at (2.5   , -5.25) {255} ;

    % \draw[->] (e50) -- (r40) ;
    % \draw[->] (e10) -- (r40) ;

    \node at (2, -6.5) {$765$} ;

    \node at (6, 0.7) {\parbox{3cm}{\subcaption{Filter        \hfill \; \label{countdown-filter}}}} ;
    \node at (6,-0.8) {\parbox{3cm}{\subcaption{Permutation   \hfill \; \label{countdown-permutation}}}} ;
    \node at (6,-2.3) {\parbox{3cm}{\subcaption{Operators     \hfill \; \label{countdown-operators}}}} ;
    \node at (6,-4.8) {\parbox{3cm}{\subcaption{Parenthesising\hfill \; \label{countdown-parens}}}} ;
  \end{tikzpicture}
  \caption{The components of a transformation which makes up a Countdown
    candidate solution}
\end{figure}

\section{Enumerating Combinations and Permutations}
\section{Enumerating Binary Trees}
\section{Filtering Out Invalid Expressions with the Subobject Classifier}
\section{Filtering Out Duplicates with Quotients}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../paper"
%%% End: